### 2021-07-22 writeUps/MOM.tex
reorganize

\documentclass[11pt]{article} % use larger type; default would be 10pt

\usepackage[utf8]{inputenc} % set input encoding (not needed with XeLaTeX)
\usepackage{amsmath,amssymb}
\usepackage{bm}
\usepackage{natbib}

\newcommand{\EE}{\mathbb{E}}
\newcommand{\pp}{e(\bm{X})}
\newcommand{\ppi}{e(\bm{X_i})}
\newcommand{\hpp}{\hat{e}(\bm{X})}
\newcommand{\hppi}{\hat{e}(\bm{X_i})}

\newcommand\independent{\protect\mathpalette{\protect\independenT}{\perp}}
\def\independenT#1#2{\mathrel{\rlap{$#1#2$}\mkern2mu{#1#2}}}

\title{MOM PS Estimator under Strong Monotonicity}

\begin{document}
\maketitle

Say members of the control group of an RCT have no access to the treatment, but among the treatment group there are compliers and non-compliers---i.e. one-sided non-compliance.
(Of course, this could be, e.g., compliance in two different ways, etc.)
Then there are two principal strata $S$, never-takers and compliers. 
Say the exclusion restriction doesn't hold.

\citet{feller2017principal} and \citet{dingLu} and others discuss estimating principal effects in one-sided noncompliance (which they call ``strong monotonicity'') under the assumption of principal ignorability, viz. $Y_C\independent S |\bm{X}$.
This has two problems, to my mind:
\begin{enumerate}
\item Sometimes principal ignorability is not plausible
\item If you assume the principal ignorability for $Y_T$ (``strong'' principal ignorability) then, as \citet{feller2017principal} points out, the principal effects conditional on covariates are equal to the ITT conditional on covariates---the stratum plays no role. But sometimes the role of the strata is what we're trying to estimate!
\end{enumerate}

The approach here takes a different tack, based on a different assumption which is maybe even less plausible---but which, I think, can be relaxed.

\section{Setup}
Estimating $\EE[Y_T|S]$ is straightforward, so for the remainder 
just consider the ``control'' group, $Z=0$, with $n$ (control) subjects and drop the $C$ subscript. 
The control group is a mixture of compliers and never-takers.

For subject $i$, $i=1,\dots,n$, we have:
\begin{itemize}
\item $Y_i$ observed outcome
\item $S_i \in \{0,1\}$ unobserved stratum
\item $\ppi=Pr(S_i=1|\bm{X}_i)$ take as given (like, estimated from treatment group)
\end{itemize}

Let $\mu_0=\EE[Y|S=0]$ and $\mu_1=\EE[Y|S=1]$.
The goal is to estimate $\mu_0$ and $\mu_1$.

\textbf{Assumption:}\\
\begin{equation}\label{eq:assumption}
\EE[Y_i|\bm{X}_i,S_i]=\EE[Y_i|S_i]=\mu_0\text{ or }\mu_1
\end{equation}
i.e. $Y$ is mean-independent of $\bm{X}$ conditional on $S$. 
This is problematic---why should $S$ contain all information about $Y$?

Note, it is kinda related to ``principal ignorability`` of \citet{feller2017principal} and \citet{dingLu}, $Y\independent S |\bm{X}_i$ but, obviously, different too.

\section{MOM Estimator}

First, relating $\bar{Y}$ to $\mu_0$ and $\mu_1$.

\begin{align*}
  \EE[Y_i|\bm{X}_i]&=\EE\left\{\EE[Y_i|\bm{X}_i,S_i]|\bm{X}_i\right\}\\
  &=\EE\left\{\EE[Y_i|S_i]|\bm{X}_i\right\}\\
  =\mu_1\ppi+\mu_0(1-\ppi)
\end{equation*}
Hence the first estimating equation is
\begin{equation}\label{eq:est1}
\bar{Y}=\hat{\mu}_1\overline{\pp}+\hat{\mu}_0(1-\overline{\pp})
\end{equation}

Second, 
\begin{align*}
\EE[Y_i\ppi]&=\EE\EE[Y_i\ppi|\ppi]\\
&=\EE\left\{ \ppi \EE\EE[Y_i|\ppi,S_i]\right\}\\
&=\EE\left\{ \ppi \EE\EE[Y_i|S_i]\right\}\text{ by \eqref{eq:assumption}}\\
&=\EE\left\{\mu_1\ppi^2+\mu_0(1-\ppi)\ppi\right\}
\end{align*}

Then the 2nd estimating equation is
\begin{equation*}
\overline{Y\pp}=\hat{\mu}_1\overline{\pp^2}+\hat{\mu}_0(\overline{\pp}-\overline{\pp^2})
\end{equation*}

Solving gives the estimates:
\begin{equation*}
\hat{\mu}_0=\frac{\overline{\pp}\;\overline{Y\pp}}{(\bar{Y}-1)\overline{\pp^2}+\overline{\pp}^2}
\end{equation*}
and
\begin{equation*}
\hat{\mu}_1=\frac{\bar{Y}-(1-\overline{\pp})\hat{\mu}_0}{\overline{\pp}}
\end{equation*}

\section{Relaxing \eqref{eq:assumption} with regression}
There's no particular reason to assume \eqref{eq:assumption} and it seems like it would typically be pretty implausible.

But say you believe the model
\begin{equation}\label{eq:regression}
Y_i=\bm{x}_i'\bm{\beta}+\gamma S_i+\epsilon_i
\end{equation}
 Then instead of \eqref{eq:assumption} we could assume something like
\begin{equation*}
\EE[Y_i-\bm{x}_i'\bm{\beta}|\bm{X}_i,S_i]=\EE[Y_i-\bm{x}_i'\bm{\beta}|S_i]
\end{equation*}

in other words, no interaction between $\bm{X}$ and $S$ in \eqref{eq:assumption}. 
Then, if you had an estimate for $\beta$, you could just substitute $Y_i-\bm{x}_i'\bm{\beta}$ for $Y_i$ in the estimates. 
Alternatively, you could use a stacked estimating equation approach and estimate it all together.

\bibliographystyle{unsrtnat}
\bibliography{MOM}

\end{document}