### 2021-11-08 writeUps/regression.tex
putting some things down in writing

\documentclass[11pt]{article} % use larger type; default would be 10pt

\usepackage[utf8]{inputenc} % set input encoding (not needed with XeLaTeX)
\usepackage{amsmath,amssymb}
\usepackage{bm}
\usepackage{natbib}
\usepackage{hyperref}
\usepackage{graphicx}

\newcommand{\EE}{\mathbb{E}}
\newcommand{\pp}{e(\bm{X})}
\newcommand{\ppi}{e(\bm{X_i})}
\newcommand{\hpp}{\hat{e}(\bm{X})}
\newcommand{\hppi}{\hat{e}(\bm{X_i})}

\newcommand\independent{\protect\mathpalette{\protect\independenT}{\perp}}
\def\independenT#1#2{\mathrel{\rlap{$#1#2$}\mkern2mu{#1#2}}}

\title{MOM Estimator: Regression Form}

\begin{document}
\maketitle

Let $p_i=Pr(S_i=1|\bm{x}_i)$.
Then the original estimating equations were:
\begin{align*}
  Z_i=0:&\\
  &Y_i-\mu_{00}(1-p_i)+\mu_{01}p_i=Y_i-\mu_{00}-(\mu_{01}-\mu_{00})p_i\\
  p_i&Y_i-\mu_{00}p_i-(\mu_{01}-\mu_{00})p_i^2\\
  Z_i=1:&\\
        &Y_i-\mu_{10}-S_i(\mu_{11}-\mu_{10})\\
  S_i&Y_i-S_i\mu_{10}-S_i(\mu_{11}-\mu_{10})=S_iY_i-S_i\mu_{11}\\
\end{align*}

Let
\begin{equation}
  r_i=
  \begin{cases}
    p_i&\text{ if }Z_i=0\\
    S_i&\text{ if }Z_i=1
  \end{cases}
\end{equation}

Note that since $r_i$ depends on $Z_i$, we can replace both $p$ and $S$ in the estimating equations with $r$, without changing them.

\begin{align*}
  Z_i=0:&\\
  &Y_i-\mu_{00}(1-r_i)+\mu_{01}r_i=Y_i-\mu_{00}-(\mu_{01}-\mu_{00})r_i\\
  r_i&Y_i-\mu_{00}r_i-(\mu_{01}-\mu_{00})r_i^2\\
  Z_i=1:&\\
        &Y_i-\mu_{10}-r_i(\mu_{11}-\mu_{10})\\
  r_i&Y_i-r_i\mu_{10}-r_i(\mu_{11}-\mu_{10})=r_iY_i-r_i\mu_{11}\\
\end{align*}

Then we can combine these into four equations that look more like OLS. The first:

\begin{align*}
  &Y_i-(1-Z_i)(\mu_{00}-(\mu_{01}-\mu_{00})r_i)-Z_i(\mu_{10}-r_i(\mu_{11}-\mu_{10}))\\
  =& Y_i-\mu_{00}-(\mu_{01}-\mu_{00})r_i-Z_i(\mu_{10}-\mu_{00})-r_iZ_i
\end{align*}
and the 2nd:
\begin{align*}
  &r_iY_i-(1-Z_i)\left[\mu_{00}r_i-(\mu_{01}-\mu_{00})r_i^2\right]-Z_ir_i\mu_{11}\\
  =&r_iY_i-\mu_{00}r_i-(\mu_{01}-\mu_{00})r_i^2-Z_ir_i\left[\mu_{11}-\mu_{00}-r_i(\mu_{01}-\mu_{00})\right]\\
\end{align*}

The 3rd and 4th are the same as above:
\begin{align*}
  Z_i&Y_i-Z_i\mu_{10}-Z_ir_i(\mu_{11}-\mu_{10})\text{ and }\\
  Z_ir_i&Y_i-Z_ir_i\mu_{11}\\
\end{align*}

These are the estimating equations for the OLS model:
\begin{equation}\label{eq:regression0}
  Y_i=\beta_0+\beta_1r_i+\beta_2Z_i+\beta_3Z_ir_i+\epsilon_i
\end{equation}
with
\begin{align*}
  \beta_0&=\mu_{00}\\
  \beta_1&=\mu_{01}-\mu_{00}\\
  \beta_2&=\mu_{10}-\mu_{00}\\
  \beta_3&=\mu_{11}-\mu_{10}-(\mu_{01}-\mu_{00})
\end{align*}

\section{Missing $S$ in Treatment Group}
When $S$ is missing at random (conditional on $\bm{x}$) for some observations in the treatment group,
and $\mathbb{E}[Y|Z=1,S,p]=\mathbb{E}[Y|Z=1,S]$,
a similar mixed model will work for estimating $\mu_{10}$ and $\mu_{11}$.
\begin{align*}
  Z_i=1\text{ and }S_i\text{ Missing}:&\\
  &Y_i-\mu_{10}(1-p_i)+\mu_{11}p_i=Y_i-\mu_{10}-(\mu_{11}-\mu_{10})p_i\\
  p_i&Y_i-\mu_{10}p_i-(\mu_{11}-\mu_{10})p_i^2\\
  Z_i=1\text{ and }S_i\text{ Observed}:&\\
        &Y_i-\mu_{10}-S_i(\mu_{11}-\mu_{10})\\
  S_i&Y_i-S_i\mu_{10}-S_i(\mu_{11}-\mu_{10})=S_iY_i-S_i\mu_{11}\\
    Z_i=0:&\\
  &Y_i-\mu_{00}(1-p_i)+\mu_{01}p_i=Y_i-\mu_{00}-(\mu_{01}-\mu_{00})p_i\\
  p_i&Y_i-\mu_{00}p_i-(\mu_{01}-\mu_{00})p_i^2\\
\end{align*}

This works because
\begin{equation*}
  \mathbb{E}[Y|p]=\mathbb{E}\left[\mathbb{E}[Y|p,S]|p\right]=\mathbb{E}\left[\mathbb{E}[Y|S]|p\right]=\mu_{10}(1-p)+\mu_{11}p
\end{equation*}

Then re-define

\begin{equation}
  r_i=
  \begin{cases}
    p_i&\text{ if }Z_i=0\text{ or }Z_i=1\text{ and }S_i\text{ missing}\\
    S_i&\text{ if }Z_i=1\text{ and }S_i\text{ observed}
  \end{cases}
\end{equation}
Then the rest of the logic holds, and the regression (\ref{eq:regression0}) can be used to estimate $\bm{\mu}$.

\section{Sandwich Standard Errors}

So we're estimating principal effects in two stages.
First, fit the model
\begin{equation}\label{eq:psMod}
  Pr(S_{Ti}=1|\bm{\tilde{x}}_i)=f(\bm{\alpha}'\bm{\tilde{x}}_i)
\end{equation}
for some inverse link function $f(\cdot)$, where $\bm{\tilde{x}}_i=[1,\bm{x}_i]$, using (observed) values from the treatment group, and estimating $\hat{\alpha}$.
Then let
\begin{equation}\label{eq:ps}
  \hat{p}_i=f(\bm{\hat{\alpha}}'\bm{\tilde{x}}_i)
\end{equation}
for all subjects in the experiment.

Finally, fit model
\begin{equation}\label{eq:regression}
  Y_i=\beta_0+\beta_1r_i+\beta_2Z_i+\beta_3Z_ir_i+\bm{\gamma}^t\bm{x}_i+\epsilon_i
\end{equation}
to estimate $\bm{\beta}$ and hence principal effects.
[Note that \eqref{eq:regression} is just \eqref{eq:regression0} with $\bm{\gamma}^t\bm{x}_i$ added on the end].
Let $\bm{\beta}=[\beta_0,\beta_1,\beta_2,\beta_3,\bm{\gamma}^t]^t$.

The standard error matrix for $\bm{\hat{\beta}}$ can be estimated as:
\begin{equation}\label{eq:sandwich}
  \widehat{var}(\bm{\hat{\beta}})=A^{-1}BA^{-t}
\end{equation}
Where
\begin{equation}\label{eq:Amat}
  A=\sum_i\frac{\partial}{\partial [\alpha,\beta]^t} \Psi(Y_i;\alpha,\beta,\bm{x}_i)
\end{equation}
and
\begin{equation}\label{eq:Bmat}
  B=\sum_i \Psi(Y_i;\alpha,\beta,\bm{x}_i)\Psi(Y_i;\alpha,\beta,\bm{x}_i)^t
\end{equation}
where $\Psi(\cdot)$ is the estimating equations.

Following \citet[][p. 373]{carroletal06}, we %separate the estimating equations $\Psi$ into $\phi$ and $\psi$, estimating equations for models \eqref{
can decompose the matrices into diagonal elements $A_{1,1}$ and $A_{2,2}$ and $B_{1,1}$ and $B_{2,2}$ that pertain to the parameter sets $\alpha$ and $\beta$ and the estimating equations for models \eqref{eq:psMod} and  \eqref{eq:regression}, respectively, and $A_{21}$, and $B_{12}=B_{21}^t$, which capture to the dependence of model \eqref{eq:regression} on the parameters $\alpha$ from \eqref{psMod} and the covariance between the estimating equations of the two models. The sub-matrix $A_{12}=0$, since \eqref{eq:psMod} does not depend on $\beta$.

The diagonal matrices $A_{1,1}$ and $A_{2,2}$ and $B_{1,1}$ and $B_{2,2}$ are all the typical ``bread'' and ``meat'' matrices from M-estimation of generalized linear models and OLS. In practice, we use the estimates from the \texttt{sandwich} package in \texttt{R}, adjusted in two ways: first, the function \texttt{bread} actually gives $A^{-1}$, not $A$; second, we must pay careful attention to sample sizes, since the sample size for \eqref{eq:psMod} includes only treated observations (with observed $S$) and \eqref{eq:regression} contains all observations.

\subsection{$A_{21}$ Matrix}
The estimating equations for the regression \eqref{eq:regression} are
\begin{equation}\label{eq:eeOLS}
  \psi(Y_i,\bm{x}_i,\beta,\alpha)=X_iY_i-X_iX_i^t\beta
\end{equation}
Where $X_i=[1,r_i,Z_i,r_iZ_i,\bm{x}_i^t]^t$.
These depend on $\alpha$ when $r_i=p_i$, i.e. when $Z_i=0$ or $S_i$ is missing.
Then note that, following \eqref{eq:ps}, and letting $\eta_i=\alpha^t\bm{\tilde{x}}_i$
\begin{equation}\label{eq:derivP}
  \frac{\partial p_i}{\partial \alpha^t}=f'(\eta_i)\frac{\partial \eta_i}{\partial \alpha^t}=f'(\eta_i)\bm{\tilde{x}}_i^t
\end{equation}
and that
\begin{equation}
  \frac{\partial p_i^2}{\partial \alpha^t}=2p\frac{\partial p_i}{\partial \alpha^t}=2f(\eta)f'(\eta_i)\bm{\tilde{x}}_i^t
\end{equation}

Then if $r_i=p_i$,
\begin{align*}
  \frac{\partial}{\partial \alpha^t}&\psi(Y_i,\bm{x}_i,\beta,\alpha)=\\
  \left[\begin{array}{c}
          -\beta_1\frac{\partial p_i}{\partial \alpha^t}\\
          (Y_i-\beta_0-\beta_2Z_i-\bm{\gamma}^t\bm{x}_{i})\partial p/\partial\alpha^t-(\beta_1+\beta_3Z_i)\partial p^2/\partial \alpha^t\\
          -\beta_1\frac{\partial p_i}{\partial \alpha^t}Z_i\\
          Z_i\left[(Y_i-\beta_0-\beta_2Z_i-\bm{\gamma}^t\bm{x}_{i})\partial p/\partial\alpha^t-(\beta_1+\beta_3Z_i)\partial p^2/\partial \alpha^t\right]\\
          -\beta_1\frac{\partial p_i}{\partial \alpha^t}\bm{x}_i
        \end{array}\right]\\
  =& \left[\begin{array}{c}
          -\beta_1f'(\eta_i)\bm{\tilde{x}}_i^t\\
          \left[Y_i-X_i^t\beta-p_i(\beta_1+\beta_3Z_i)\right]f'(\eta_i)\bm{\tilde{x}}_i\\
          -Z_i\beta_1f'(\eta_i)\bm{\tilde{x}}_i^t\\
          Z_i\left[Y_i-X_i^t\beta-p_i(\beta_1+\beta_3)\right]f'(\eta_i)\bm{\tilde{x}}_i\\
          -\beta_1f'(\eta_i)\bm{x}_i\bm{x}_i^t
           \end{array}\right]
\end{align*}
Note that unless there is missing $S$ in the treatment group, $Z_i=0$ whenever $r_i=p_i$, so the 3rd and 4th lines of the matrix are all 0.

If $r_i=S_i$, $\frac{\partial}{\partial \alpha^t}&\psi(Y_i,\bm{x}_i,\beta,\alpha)=0$.

\end{document}