### 2022-09-30 writeUps/rmseTab.tex
update paper

\begin{tabular}[t]{lllrrrrrrrrr}
\toprule
\multicolumn{3}{c}{ } & \multicolumn{3}{c}{$\alpha=0$} & \multicolumn{3}{c}{$\alpha=0.2$} & \multicolumn{3}{c}{$\alpha=0.5$} \\
\cmidrule(l{3pt}r{3pt}){4-6} \cmidrule(l{3pt}r{3pt}){7-9} \cmidrule(l{3pt}r{3pt}){10-12}
\makecell[l]{Residual\\Dist.} & \makecell[c]{X:Z\\Int.?} & \makecell[r]{X:S\\Int.?} & GEEPERs & Mix. & PSW & GEEPERs & Mix. & PSW & GEEPERs & Mix. & PSW\\
\midrule
 & No & No & 2.07 & 0.19 & 0.17 & 0.55 & 0.20 & 0.20 & 0.19 & 0.16 & 0.24\\

 & Yes & No & 9.34 & 0.34 & 0.16 & 1.06 & 0.33 & 0.21 & 0.23 & 0.18 & 0.23\\

 & No & Yes & 2.56 & 0.19 & 0.16 & 0.51 & 0.22 & 0.21 & 0.19 & 0.15 & 0.24\\

\multirow{-4}{*}{\raggedright\arraybackslash Normal} & Yes & Yes & 7.16 & 0.34 & 0.17 & 1.81 & 0.33 & 0.20 & 0.23 & 0.20 & 0.24\\
\cmidrule{1-12}
 & No & No & 4.19 & 0.56 & 0.16 & 0.60 & 0.50 & 0.20 & 0.18 & 0.31 & 0.24\\

 & Yes & No & 6.99 & 0.60 & 0.17 & 0.91 & 0.57 & 0.20 & 0.22 & 0.35 & 0.24\\

 & No & Yes & 1.83 & 0.52 & 0.17 & 0.45 & 0.46 & 0.20 & 0.19 & 0.31 & 0.23\\

\multirow{-4}{*}{\raggedright\arraybackslash Uniform} & Yes & Yes & 7.55 & 0.60 & 0.18 & 1.28 & 0.56 & 0.20 & 0.24 & 0.35 & 0.24\\
\bottomrule
\end{tabular}