### 2022-12-19 writeUps/outcomeRegAppendix.tex
application results appendix

\begin{table}
\begin{center}
\begin{tabular}{l c c c c}
\hline
 & psModel & BAU & FH2T & DragonBox \\
\hline
(Intercept)                                        & $-2.14$       & $1.73$      & $2.83^{***}$ & $2.51^{**}$  \\
                                                   & $(1.26)$      & $(0.92)$    & $(0.72)$     & $(0.92)$     \\
Scale.Score5                                       & $-0.59^{**}$  & $0.21$      & $0.45^{***}$ & $0.05$       \\
                                                   & $(0.19)$      & $(0.13)$    & $(0.10)$     & $(0.13)$     \\
ESOL1                                              & $-0.83$       & $0.09$      & $-0.46$      & $-0.49$      \\
                                                   & $(0.64)$      & $(0.36)$    & $(0.27)$     & $(0.35)$     \\
IEP1                                               & $-1.34^{**}$  & $0.15$      & $0.23$       & $0.10$       \\
                                                   & $(0.46)$      & $(0.29)$    & $(0.25)$     & $(0.28)$     \\
pre.total\_time\_on\_tasks                         & $-0.45^{**}$  & $-0.13$     & $0.08$       & $-0.00$      \\
                                                   & $(0.15)$      & $(0.09)$    & $(0.07)$     & $(0.10)$     \\
pre\_MSE\_total\_score                             & $-0.18$       & $0.20^{*}$  & $0.12$       & $0.35^{***}$ \\
                                                   & $(0.13)$      & $(0.08)$    & $(0.07)$     & $(0.08)$     \\
fullYear5TRUE                                      & $1.40^{*}$    & $-0.13$     & $-0.10$      & $0.06$       \\
                                                   & $(0.58)$      & $(0.33)$    & $(0.29)$     & $(0.34)$     \\
pre\_MA\_total\_scoreNATRUE                        & $-3.20^{***}$ &             &              &              \\
                                                   & $(0.71)$      &             &              &              \\
GenderM                                            & $0.17$        & $-0.19$     & $-0.34^{**}$ & $-0.27$      \\
                                                   & $(0.25)$      & $(0.15)$    & $(0.13)$     & $(0.16)$     \\
raceEthHispanic/Latino                             & $-0.56$       & $0.28$      & $0.33$       & $0.23$       \\
                                                   & $(0.46)$      & $(0.29)$    & $(0.21)$     & $(0.28)$     \\
raceEthAsian                                       & $-0.43$       & $0.66^{*}$  & $0.62^{**}$  & $0.82^{**}$  \\
                                                   & $(0.42)$      & $(0.28)$    & $(0.22)$     & $(0.30)$     \\
raceEthOther                                       & $-0.13$       & $0.45$      & $0.21$       & $0.10$       \\
                                                   & $(0.49)$      & $(0.29)$    & $(0.23)$     & $(0.30)$     \\
GIFTED1                                            & $-0.24$       & $0.66^{**}$ & $0.43^{*}$   & $0.34$       \\
                                                   & $(0.37)$      & $(0.24)$    & $(0.19)$     & $(0.26)$     \\
poly(pre\_PS\_tasks\_total\_score, 2, raw = TRUE)1 & $0.01$        &             &              &              \\
                                                   & $(0.19)$      &             &              &              \\
poly(pre\_PS\_tasks\_total\_score, 2, raw = TRUE)2 & $-0.41^{**}$  &             &              &              \\
                                                   & $(0.13)$      &             &              &              \\
Z                                                  &               & $0.29$      & $-0.17$      & $-0.20$      \\
                                                   &               & $(0.28)$    & $(0.22)$     & $(0.27)$     \\
Sp                                                 &               & $0.03$      & $-0.22$      & $0.04$       \\
                                                   &               & $(0.52)$    & $(0.41)$     & $(0.52)$     \\
pre\_PS\_tasks\_total\_score                       &               & $0.34^{**}$ & $0.37^{***}$ & $0.23$       \\
                                                   &               & $(0.12)$    & $(0.09)$     & $(0.12)$     \\
pre.total\_math\_score                             &               & $0.14^{**}$ & $0.24^{***}$ & $0.15^{**}$  \\
                                                   &               & $(0.05)$    & $(0.04)$     & $(0.05)$     \\
pre.total\_math\_scoreNATRUE                       &               & $-0.35$     & $0.18$       & $-0.44$      \\
                                                   &               & $(0.45)$    & $(0.36)$     & $(0.43)$     \\
Z:Sp                                               &               & $-0.14$     & $0.17$       & $-0.17$      \\
                                                   &               & $(0.49)$    & $(0.39)$     & $(0.49)$     \\
\hline
AIC                                                & $447.55$      & $$          & $$           & $$           \\
BIC                                                & $539.47$      & $$          & $$           & $$           \\
Log Likelihood                                     & $-200.78$     & $$          & $$           & $$           \\
Deviance                                           & $401.55$      & $$          & $$           & $$           \\
Num. obs.                                          & $402$         & $787$       & $1193$       & $771$        \\
R$^2$                                              & $$            & $0.60$      & $0.60$       & $0.63$       \\
Adj. R$^2$                                         & $$            & $0.52$      & $0.55$       & $0.54$       \\
\hline
\multicolumn{5}{l}{\scriptsize{$^{***}p<0.001$; $^{**}p<0.01$; $^{*}p<0.05$}}
\end{tabular}
\caption{Statistical models}
\label{table:coefficients}
\end{center}
\end{table}