### 2024-12-29 writeUps/regression.tex
update

\documentclass[11pt]{article} % use larger type; default would be 10pt

\usepackage[utf8]{inputenc} % set input encoding (not needed with XeLaTeX)
\usepackage{amsmath,amssymb}
\usepackage{bm}
\usepackage{natbib}
\usepackage{hyperref}
\usepackage{graphicx}

\newcommand{\EE}{\mathbb{E}}
\newcommand{\pp}{e(\bm{X})}
\newcommand{\ppi}{e(\bm{X_i})}
\newcommand{\hpp}{\hat{e}(\bm{X})}
\newcommand{\hppi}{\hat{e}(\bm{X_i})}

\newcommand\independent{\protect\mathpalette{\protect\independenT}{\perp}}
\def\independenT#1#2{\mathrel{\rlap{$#1#2$}\mkern2mu{#1#2}}}

\title{MOM Estimator: Regression Form}

\begin{document}
\maketitle

Let $p_i=Pr(S_i=1|\bm{w}_i)$, where $\bm{w}_i$ is a vector of covariates that (putatively) predict $S_i$.
(Really this should be $S_{Ti}$, cuz we're doing principal stratification; we'll drop the $T$ subscript for clarity.)

Then the original estimating equations were:
\begin{align*}
  Z_i=0:&\\
  &Y_i-\mu_{00}(1-p_i)+\mu_{01}p_i=Y_i-\mu_{00}-(\mu_{01}-\mu_{00})p_i\\
  p_i&Y_i-\mu_{00}p_i-(\mu_{01}-\mu_{00})p_i^2\\
  Z_i=1:&\\
        &Y_i-\mu_{10}-S_i(\mu_{11}-\mu_{10})\\
  S_i&Y_i-S_i\mu_{10}-S_i(\mu_{11}-\mu_{10})=S_iY_i-S_i\mu_{11}\\
\end{align*}

Let
\begin{equation}
  r_i=
  \begin{cases}
    p_i&\text{ if }Z_i=0\\
    S_i&\text{ if }Z_i=1
  \end{cases}
\end{equation}

Note that since $r_i$ depends on $Z_i$, we can replace both $p$ and $S$ in the estimating equations with $r$, without changing them.

\begin{align*}
  Z_i=0:&\\
  &Y_i-\mu_{00}(1-r_i)+\mu_{01}r_i=Y_i-\mu_{00}-(\mu_{01}-\mu_{00})r_i\\
  r_i&Y_i-\mu_{00}r_i-(\mu_{01}-\mu_{00})r_i^2\\
  Z_i=1:&\\
        &Y_i-\mu_{10}-r_i(\mu_{11}-\mu_{10})\\
  r_i&Y_i-r_i\mu_{10}-r_i(\mu_{11}-\mu_{10})=r_iY_i-r_i\mu_{11}\\
\end{align*}

Then we can combine these into four equations that look more like OLS. The first:

\begin{align*}
  &Y_i-(1-Z_i)(\mu_{00}-(\mu_{01}-\mu_{00})r_i)-Z_i(\mu_{10}-r_i(\mu_{11}-\mu_{10}))\\
  =& Y_i-\mu_{00}-(\mu_{01}-\mu_{00})r_i-Z_i(\mu_{10}-\mu_{00})-r_iZ_i
\end{align*}
and the 2nd:
\begin{align*}
  &r_iY_i-(1-Z_i)\left[\mu_{00}r_i-(\mu_{01}-\mu_{00})r_i^2\right]-Z_ir_i\mu_{11}\\
  =&r_iY_i-\mu_{00}r_i-(\mu_{01}-\mu_{00})r_i^2-Z_ir_i\left[\mu_{11}-\mu_{00}-r_i(\mu_{01}-\mu_{00})\right]\\
\end{align*}

The 3rd and 4th are the same as above:
\begin{align*}
  Z_i&Y_i-Z_i\mu_{10}-Z_ir_i(\mu_{11}-\mu_{10})\text{ and }\\
  Z_ir_i&Y_i-Z_ir_i\mu_{11}\\
\end{align*}

These are the estimating equations for the OLS model:
\begin{equation}\label{eq:regression0}
  Y_i=\beta_0+\beta_1r_i+\beta_2Z_i+\beta_3Z_ir_i+\epsilon_i
\end{equation}
with
\begin{align*}
  \beta_0&=\mu_{00}\\
  \beta_1&=\mu_{01}-\mu_{00}\\
  \beta_2&=\mu_{10}-\mu_{00}\\
  \beta_3&=\mu_{11}-\mu_{10}-(\mu_{01}-\mu_{00})
\end{align*}

\section{Missing $S$ in Treatment Group}
When $S$ is missing at random (conditional on $\bm{w}$) for some observations in the treatment group,
and $\mathbb{E}[Y|Z=1,S,p]=\mathbb{E}[Y|Z=1,S]$,
a similar mixed model will work for estimating $\mu_{10}$ and $\mu_{11}$.
\begin{align*}
  Z_i=1\text{ and }S_i\text{ Missing}:&\\
  &Y_i-\mu_{10}(1-p_i)+\mu_{11}p_i=Y_i-\mu_{10}-(\mu_{11}-\mu_{10})p_i\\
  p_i&Y_i-\mu_{10}p_i-(\mu_{11}-\mu_{10})p_i^2\\
  Z_i=1\text{ and }S_i\text{ Observed}:&\\
        &Y_i-\mu_{10}-S_i(\mu_{11}-\mu_{10})\\
  S_i&Y_i-S_i\mu_{10}-S_i(\mu_{11}-\mu_{10})=S_iY_i-S_i\mu_{11}\\
    Z_i=0:&\\
  &Y_i-\mu_{00}(1-p_i)+\mu_{01}p_i=Y_i-\mu_{00}-(\mu_{01}-\mu_{00})p_i\\
  p_i&Y_i-\mu_{00}p_i-(\mu_{01}-\mu_{00})p_i^2\\
\end{align*}

This works because
\begin{equation*}
  \mathbb{E}[Y|p]=\mathbb{E}\left[\mathbb{E}[Y|p,S]|p\right]=\mathbb{E}\left[\mathbb{E}[Y|S]|p\right]=\mu_{10}(1-p)+\mu_{11}p
\end{equation*}

Then re-define

\begin{equation}
  r_i=
  \begin{cases}
    p_i&\text{ if }Z_i=0\text{ or }Z_i=1\text{ and }S_i\text{ missing}\\
    S_i&\text{ if }Z_i=1\text{ and }S_i\text{ observed}
  \end{cases}
\end{equation}
Then the rest of the logic holds, and the regression (\ref{eq:regression0}) can be used to estimate $\bm{\mu}$.

\section{Sandwich Standard Errors}

So we're estimating principal effects in two stages.
First, fit the model
\begin{equation}\label{eq:psMod}
  Pr(S_{Ti}=1|\bm{\tilde{w}}_i)=f(\bm{\alpha}'\bm{\tilde{w}}_i)
\end{equation}
for some inverse link function $f(\cdot)$, where $\bm{\tilde{x}}_i=[1,\bm{w}_i]$, using (observed) values from the treatment group, and estimating $\hat{\alpha}$.
Then let
\begin{equation}\label{eq:ps}
  \hat{p}_i=f(\bm{\hat{\alpha}}'\bm{\tilde{w}}_i)
\end{equation}
for all subjects in the experiment.

Finally, fit model
\begin{equation}\label{eq:regression}
  Y_i=\beta_0+\beta_1r_i+\beta_2Z_i+\beta_3Z_ir_i+\bm{\gamma}^t\bm{x}_i+\epsilon_i
\end{equation}
to estimate $\bm{\beta}$ and hence principal effects, where $\bm{x}_i$ is a set of covariates predictive of $Y$ within principal strata.
[Note that \eqref{eq:regression} is just \eqref{eq:regression0} with $\bm{\gamma}^t\bm{x}_i$ added on the end].
$\bm{x}_i$ should, generally, contain $\bm{w}_i$; the two may also be equivalent.
Let $\bm{\beta}=[\beta_0,\beta_1,\beta_2,\beta_3,\bm{\gamma}^t]^t$.

The standard error matrix for $\bm{\hat{\beta}}$ can be estimated as:
\begin{equation}\label{eq:sandwich}
  \widehat{var}(\bm{\hat{\beta}})=A^{-1}BA^{-t}
\end{equation}
Where
\begin{equation}\label{eq:Amat}
  A=\sum_i\frac{\partial}{\partial [\alpha,\beta]^t} \Psi(Y_i;\alpha,\beta,\bm{x}_i,\bm{w}_i)
\end{equation}
and
\begin{equation}\label{eq:Bmat}
  B=\sum_i \Psi(Y_i;\alpha,\beta,\bm{x}_i)\Psi(Y_i;\alpha,\beta,\bm{x}_i,\bm{w}_i)^t
\end{equation}
where $\Psi(\cdot)$ is the estimating equations.

Following \citet[][p. 373]{carroletal06}, we %separate the estimating equations $\Psi$ into $\phi$ and $\psi$, estimating equations for models \eqref{
can decompose the matrices into diagonal elements $A_{1,1}$ and $A_{2,2}$ and $B_{1,1}$ and $B_{2,2}$ that pertain to the parameter sets $\alpha$ and $\beta$ and the estimating equations for models \eqref{eq:psMod} and  \eqref{eq:regression}, respectively, and $A_{21}$, and $B_{12}=B_{21}^t$, which capture to the dependence of model \eqref{eq:regression} on the parameters $\alpha$ from \eqref{psMod} and the covariance between the estimating equations of the two models. The sub-matrix $A_{12}=0$, since \eqref{eq:psMod} does not depend on $\beta$.

The diagonal matrices $A_{1,1}$ and $A_{2,2}$ and $B_{1,1}$ and $B_{2,2}$ are all the typical ``bread'' and ``meat'' matrices from M-estimation of generalized linear models and OLS. In practice, we use the estimates from the \texttt{sandwich} package in \texttt{R}, adjusted in two ways: first, the function \texttt{bread} actually gives $A^{-1}$, not $A$; second, we must pay careful attention to sample sizes, since the sample size for \eqref{eq:psMod} includes only treated observations (with observed $S$) and \eqref{eq:regression} contains all observations.

\subsection{$A_{21}$ Matrix}
The estimating equations for the regression \eqref{eq:regression} are
\begin{equation}\label{eq:eeOLS}
  \psi(Y_i,\bm{x}_i,\bm{w}_i,\beta,\alpha)=X_iY_i-X_iX_i^t\beta
\end{equation}
Where $X_i=[1,r_i,Z_i,r_iZ_i,\bm{x}_i^t]^t$.
In other words,
\begin{align*}
  \psi(Y_i,&\bm{x}_i,\bm{w}_i,\beta,\alpha)=\\
  &\left\{
  \begin{array}{l}
    Y_i-\left(\beta_0+\beta_1r_i+\beta_2Z_i+\beta_3Z_ir_i+\bm{\gamma}'\bm{x}_i\right)\\
    r_iY_i-r_i\left(\beta_0+\beta_2Z_i+\bm{\gamma}'\bm{x}_i\right)-r_i^2\left(\beta_1+\beta_3Z_i\right)\\
    Z_iY_i-Z_i\left(\beta_0+\beta_1r_i+\beta_2Z_i+\beta_3Z_ir_i+\bm{\gamma}'\bm{x}_i\right)\\
    Z_ir_iY_i-Z_ir_i\left(\beta_0+\beta_2Z_i+\bm{\gamma}'\bm{x}_i\right)-Z_ir_i^2\left(\beta_1+\beta_3Z_i\right)\\
    \bm{x}_iY_i-\bm{x}_i\left(\beta_0+\beta_1r_i+\beta_2Z_i+\beta_3Z_ir_i+\bm{\gamma}'\bm{x}_i\right)
  \end{array}
  \right\}
\end{align*}
(noting that $Z^2=Z$).
These depend on $\alpha$ when $r_i=p_i$, i.e. when $Z_i=0$ or $S_i$ is missing.
Then note that, following \eqref{eq:ps}, and letting $\eta_i=\alpha^t\bm{\tilde{w}}_i$
\begin{equation}\label{eq:derivP}
  \frac{\partial p_i}{\partial \alpha^t}=f'(\eta_i)\frac{\partial \eta_i}{\partial \alpha^t}=f'(\eta_i)\bm{\tilde{w}}_i^t
\end{equation}
and that
\begin{equation}
  \frac{\partial p_i^2}{\partial \alpha^t}=2p_i\frac{\partial p_i}{\partial \alpha^t}=2f(\eta)f'(\eta_i)\bm{\tilde{w}}_i^t=2p_if'(\eta_i)\bm{\tilde{w}}_i^t
\end{equation}

Then if $r_i=p_i$,
\begin{align*}
  \frac{\partial}{\partial \alpha^t}&\psi(Y_i,\bm{x}_i,\beta,\alpha)=\\
  &\left[\begin{array}{c}
          -(\beta_1+\beta_3Z_i)\frac{\partial p_i}{\partial \alpha^t}\\
          (Y_i-\beta_0-\beta_2Z_i-\bm{\gamma}^t\bm{x}_{i})\partial p/\partial\alpha^t-(\beta_1+\beta_3Z_i)\partial p^2/\partial \alpha^t\\
          -(\beta_1+\beta_3Z_i)\frac{\partial p_i}{\partial \alpha^t}Z_i\\
          Z_i\left[(Y_i-\beta_0-\beta_2Z_i-\bm{\gamma}^t\bm{x}_{i})\partial p/\partial\alpha^t-(\beta_1+\beta_3Z_i)\partial p^2/\partial \alpha^t\right]\\
          -(\beta_1+\beta_3Z_i)\frac{\partial p_i}{\partial \alpha^t}\bm{x}_i
        \end{array}\right]\\
  =& \left[\begin{array}{c}
          -(\beta_1+\beta_3Z_i)f'(\eta_i)\bm{\tilde{w}}_i^t\\
          \left[Y_i-\beta_0-\beta_2Z_i-\bm{x}_i^t\bm{\gamma}-2p_i(\beta_1+\beta_3Z_i)\right]f'(\eta_i)\bm{\tilde{w}}_i^t\\
          -Z_i\beta_1f'(\eta_i)\bm{\tilde{w}}_i^t\\
          Z_i\left[Y_i-\beta_0-\beta_2Z_i-\bm{x}_i^t\bm{\gamma}-2p_i(\beta_1+\beta_3)\right]f'(\eta_i)\bm{\tilde{w}}_i^t\\
          -(\beta_1+\beta_3Z_i)f'(\eta_i)\bm{x}_i\bm{\tilde{w}}_i^t
    \end{array}\right]
\end{align*}

When $S$ is observed for all members of the treatment group, $Z_i=0$ whenever $r_i=p_i$, so the latter expression reduces to
\begin{align*}
  \frac{\partial}{\partial \alpha^t}&\psi(Y_i,\bm{x}_i,\beta,\alpha)=\\
  & \left[\begin{array}{c}
          -(\beta_1)f'(\eta_i)\bm{\tilde{w}}_i^t\\
          \left[Y_i-X_i^t\beta-2p_i(\beta_1)\right]f'(\eta_i)\bm{\tilde{w}}_i^t\\
          0\\
          0\\
          -\beta_1f'(\eta_i)\bm{x}_i\bm{\tilde{w}}_i^t
    \end{array}\right]
\end{align*}

If $r_i=S_i$, $\frac{\partial}{\partial \alpha^t}\psi(Y_i,\bm{x}_i,\beta,\alpha)=0$.

\section{Sample Sizes}

I am a bit confused over this one.
In Sefanski and Boos, the sums that define $A$ and $B$ are divided by $n$, the sample size.
In Carrol et al., the sums are not divided by anything, but the variance formula has an $n^{-1}$ in it, $A^{-1}BA^{-t}/n$.
The dilemma is that the sample size for \eqref{eq:psMod} is $\approx n/2$ and the sample size for \eqref{eq:regression} is $n$.

My intuition is based on the fact that $A$ is supposed to estimate $\mathbb{E}[-d\Psi/d\theta]$ and $B$ is supposed to estimate $\mathbb{E}[\Psi\Psi^T]$.
Hence, $A_{11}$ and $B_{11}$ should be divided by the sample size for \eqref{eq:psMod}, and $A_{22}$ and $B_{22}$ should be divided by $n$.
What about $A_{21}$ and $B_{21}$? Well, the covariance between the estimating equations of the two models should account for the fact that roughly half of the estimating equations for the outcome model's sample is uncorrelated with the estimating equations for the principal score model, and the other half is not dependent on its parameters---i.e., there are zeroes in the sum. Dividing by $n$ treats those zeros as informative, which seems correct to me.

On the other hand, it turns out that if you don't divide anything by anything---i.e. there's no $n$ or $n/2$ in any denominators---you reproduce the standard errors from \texttt{geex} almost exactly.

In any event, it doesn't seem to matter much in practice, and standard errors computed either way, dividing by sample sizes or not, are all almost unbiased, in the scenarios I've simulated. So rn I'm not sure.

\end{document}